---
params:
  pid: "Patient"
  sid: "Sample"
  mid: ""
  date: "`r Sys.time()`"
  marker.channel: 1
  alb.channel: 2
  dx.alb: ""
  fu.alb: ""
  mnc.alb: ""
  h2o.alb: ""
  dx.marker: ""
  mnc.marker: ""
  h2o.marker: ""
  fu.marker: ""
  dx.baseline: NA
  mnc.alb.dilutionX: 200
  mnc.alb.concentration: ""
  show_outliers: TRUE
  run_by: ""
  run_date: ""
  reported_by: ""
  runType: ""
  verifier: ""
  is_manual_threshold: FALSE
  manual_threshold: 0
  qc.concentration: 0
title: "ddPCR laboratory results"
fontsize: 10pt
geometry: margin=1.5cm
fig_width: 8 
fig_height: 6
output:
  pdf_document:
    keep_tex: false
    toc: false
    highlight: zenburn
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lastpage}
- \usepackage{xcolor}
- \usepackage{tabularx}
- \usepackage{caption}

---
\definecolor{darkgreen}{HTML}{188a06}

<!--- Define Headers and Footers --->
\fancypagestyle{plain}{%
  \renewcommand{\headrulewidth}{0pt}%
  \fancyhf{}%
  \fancyfoot[C]{\footnotesize Page \thepage\, of\, \pageref*{LastPage} `r params$sid` `r params$mid` Generated at `r params$date` by ddPCR MRD Pipeline (for cfDNA only) Version 1.0, 12 Sep 2024}
  \setlength\footskip{0pt}
}
\pagestyle{plain}
\center
# Sample `r params$sid`
\center
# Marker `r params$mid`


```{r echo=FALSE}

defaultW <- getOption("warn") 
options(warn = -1) 
library(knitr)
options(warn = defaultW)


plot1d <- function(multiwell.results, channel, ...) {
	x <- c()
	y <- c()
	mask <- c()
	seg <- c(0)
	lower <- c()
	higher <- c()
	mids <- c()
	col <- c()
	current.x <- 0
	for (results in multiwell.results) {
		x <- c(x, current.x + (1:nrow(results$intensities)))
		
		y <- c(y, results$intensities[, channel])
		mask <- c(mask, results$mask)
		lower <- c(lower, current.x)
		mids <- c(mids, current.x + 0.5 * nrow(results$intensities))
		current.x <- current.x + nrow(results$intensities)
		higher <- c(higher, current.x)
		col <- c(col, 1 + (results$intensities[, channel] > results$threshold))
		seg <- c(seg, current.x + 0.5)
	}
	plot(x, y, cex=1, xlab="", ylab = "intensities", axes=FALSE, pch=c(4, 16)[mask + 1], col=col, ...)
	box()
	axis(2)
	axis(1, at = mids, labels = FALSE)
	
	text(x = mids,
     y = par("usr")[3] - (max(y) - min(y))/30,
     labels = sapply(multiwell.results, "[[", "name"),
     xpd = NA,
     srt = 25,
     cex = 0.8, adj=1)
	 
	for (i in 1:length(multiwell.results)) {
		lines(c(lower[i], higher[i]), c(multiwell.results[[i]]$threshold, multiwell.results[[i]]$threshold), lty=1, col=2, lwd=2)
	}
	for (i in 1:length(seg)) {
		lines(c(seg[i], seg[i]), c(1e-6, 1e6), col="blue", lty=1)
	}
}
plot2d <- function(singlewell.results, channel, ...) {
		plot(singlewell.results$intensities[, marker.channel], singlewell.results$intensities[, alb.channel], col = (singlewell.results$intensities[, channel] > singlewell.results$threshold) + 1,
			pch=c(4, 16)[singlewell.results$mask + 1],
			main=paste(singlewell.results$name, names(singlewell.results$name)), xlab = "Channel 1", ylab="Channel 2", ...)
		if (channel == 1) {
			lines(c(singlewell.results$threshold, singlewell.results$threshold), c(-1e8, 1e8), col=2, lwd=1.5)
		} else if (channel == 2) {
			lines(c(-1e8, 1e8), c(singlewell.results$threshold, singlewell.results$threshold), col=2, lwd=1.5)
		}
}
get_info <- function(multiwell.results) {
	tab <- t(sapply(multiwell.results, function(results) {
		events <- nrow(results$intensities)
		outliers <- results$n.outliers
		used.events <- events - outliers
		threshold <- round(results$threshold)
		positive <- results$n.positive.droplets
		concentration <- round(results$concentration, digits = 3) 
		c(events, outliers, used.events, threshold, positive, concentration)
	}))
	rownames(tab) <- sapply(multiwell.results, "[[", "name")
	colnames(tab) <- c("events", "outliers", "used.events", "threshold", "positive", "concentration")
	tab
}

get_merged_info <- function(multiwell.results, sample.names=NA) {
	if (is.na(sample.names)) {
		sample.names = sapply(multiwell.results, "[[", "name")
	
	}
	tab <- get_info(multiwell.results)
	unique.names <- unique(sample.names)
	
	total.events <- tapply(tab[, "events"], sample.names, sum)[unique.names]
	total.outliers <- tapply(tab[, "outliers"], sample.names, sum)[unique.names]
	total.used.events <- tapply(tab[, "used.events"], sample.names, sum)[unique.names]
	total.positive <- tapply(tab[, "positive"], sample.names, sum)[unique.names]
	concentration <- -log(1-(total.positive/total.used.events))/0.00085
	concentration <- round(concentration[unique.names], digits = 3) 
	replicates <- tapply(tab[, "events"], sample.names, length)[unique.names]
	info <- cbind(replicates=replicates, events=total.events, outliers=total.outliers, used.events=total.used.events,positive=total.positive,concentration=concentration )
	rownames(info) <- unique.names
	info
}
```


# Marker Results

```{r fig2, fig.height=7, fig.width=12, fig.align="center", echo=FALSE}
plot1d(c(params$dx.marker, params$mnc.marker, params$h2o.marker, params$fu.marker),1, main="Intensities and annotation of marker wells")
marker.info <- get_info(c(params$dx.marker, params$mnc.marker, params$h2o.marker, params$fu.marker))
marker.merged.info <- get_merged_info(c(params$dx.marker, params$mnc.marker, params$h2o.marker, params$fu.marker))
kable(marker.info, caption = "Per well summary of marker results")
kable(marker.merged.info, caption = "Per sample summary of marker results")
```

\newpage

# Albumin Results

```{r fig1, fig.height=7, fig.width=12, fig.align="center", echo=FALSE}
plot1d(c(params$dx.alb, params$mnc.alb, params$h2o.alb, params$fu.alb), 2, main="Intensities and annotation of albumin wells")
alb.info <- get_info(c(params$dx.alb, params$mnc.alb, params$h2o.alb, params$fu.alb))
alb.merged.info <- get_merged_info(c(params$dx.alb, params$mnc.alb, params$h2o.alb, params$fu.alb))
kable(alb.info, caption = "Per well summary of Albumin results")
kable(alb.merged.info, caption = "Per sample summary of Albumin results")
```

\newpage

# Quality control


```{r qc, echo=FALSE}
## >= 8000
min.events.per.well <- min(alb.info[, "used.events"], marker.info[, "used.events"])
min.events.per.well.status <- c("\\textcolor{red}{WARNING}", "\\textcolor{darkgreen}{PASS}")[(min.events.per.well >= 8000) +  1]


## DNA quality, min FU, Alb, concentration
### >10000 good
### 2000~10000, acceptable
### < 2000 poor
fu.alb.info.perwell <- get_info(c(params$fu.alb))
fu.alb.dilution.factor.perwell <- as.numeric(gsub("x$", "", sapply(strsplit(rownames(fu.alb.info.perwell), "_", fixed=TRUE), "[[", 4)))
min.fu.alb.conc.perwell <- min(fu.alb.info.perwell[, "concentration"] * fu.alb.dilution.factor.perwell)
min.fu.alb.conc.perwell.formatted <- sprintf("%#.2f", min.fu.alb.conc.perwell)


if (min.fu.alb.conc.perwell < 2000){

	min.fu.alb.conc.perwell.status <- "\\textcolor{red}{POOR}"

	value1 <- ""
	status1 <- ""
	value2 <- ""
	status2 <- ""
	value3 <- min.fu.alb.conc.perwell.formatted
	status3 <- min.fu.alb.conc.perwell.status

} else if (min.fu.alb.conc.perwell >= 2000 && min.fu.alb.conc.perwell <= 10000) {

	min.fu.alb.conc.perwell.status  <- "\\textcolor{brown}{ACCEPTABLE}"

	value1 <- ""
	status1 <- ""
	value2 <- min.fu.alb.conc.perwell.formatted
	status2 <- min.fu.alb.conc.perwell.status
	value3 <- ""
	status3 <- ""

} else {

	min.fu.alb.conc.perwell.status  <- "\\textcolor{darkgreen}{GOOD}"

	value1 <- min.fu.alb.conc.perwell.formatted
	status1 <- min.fu.alb.conc.perwell.status
	value2 <- ""
	status2 <- ""
	value3 <- ""
	status3 <- ""

}


## controls for marker
### positive marker control >=10
dx.marker.info.perwell <- get_info(c(params$dx.marker))
min.dx.marker.con.perwell <- min(dx.marker.info.perwell[, "concentration"])
min.dx.marker.con.perwell.formatted <- sprintf("%#.2f", min.dx.marker.con.perwell)
min.dx.marker.con.perwell.status <- c("\\textcolor{red}{WARNING}", "\\textcolor{darkgreen}{PASS}")[(min.dx.marker.con.perwell >= 10) + 1]

#### Negative control just indicate value
#mnc.marker.info.perwell <- get_info(c(params$mnc.marker))
#max.mnc.marker.con.perwell <- max(mnc.marker.info.perwell[, "concentration"])
#max.mnc.marker.con.perwell.formatted <- sprintf("%#.2f", max.mnc.marker.con.perwell)
#

### NTC <= 2
h2o.marker.info.perwell <- get_info(c(params$h2o.marker))
max.h2o.marker.con.perwell <- max(h2o.marker.info.perwell[, "concentration"])
max.h2o.marker.con.perwell.formatted <- sprintf("%#.2f", max.h2o.marker.con.perwell)
max.h2o.marker.con.perwell.status <- c("\\textcolor{red}{WARNING}", "\\textcolor{darkgreen}{PASS}")[(max.h2o.marker.con.perwell <= 0.1) + 1]

## controls for albumin
### positive albumin control >= 50
mnc.alb.info.perwell <- get_info(c(params$mnc.alb))
min.mnc.alb.con.perwell <- min(mnc.alb.info.perwell[, "concentration"])
min.mnc.alb.con.perwell.formatted <- sprintf("%#.2f", min.mnc.alb.con.perwell)
min.mnc.alb.con.perwell.status <- c("\\textcolor{red}{WARNING}", "\\textcolor{darkgreen}{PASS}")[(min.mnc.alb.con.perwell >= 50) + 1]

### NTC <= 2
h2o.alb.info.perwell <- get_info(c(params$h2o.alb))
max.h2o.alb.con.perwell <- max(h2o.alb.info.perwell[, "concentration"])
max.h2o.alb.con.perwell.formatted <- sprintf("%#.2f", max.h2o.alb.con.perwell)
max.h2o.alb.con.perwell.status <- c("\\textcolor{red}{WARNING}","\\textcolor{darkgreen}{PASS}")[(max.h2o.alb.con.perwell <= 0.1) + 1]


final.qc.call <- c("\\textbf{\\textcolor{red}{WARNING}}","\\textbf{\\textcolor{darkgreen}{PASS}}")[(min.events.per.well.status == "\\textcolor{darkgreen}{PASS}" && min.dx.marker.con.perwell.status == "\\textcolor{darkgreen}{PASS}" && max.h2o.marker.con.perwell.status == "\\textcolor{darkgreen}{PASS}" && min.mnc.alb.con.perwell.status == "\\textcolor{darkgreen}{PASS}" && max.h2o.alb.con.perwell.status == "\\textcolor{darkgreen}{PASS}") + 1]




```


\captionof{table}{Quality control summary}
\vspace*{3mm}
\begin{tabular}{p{6cm}p{6cm}p{2cm}p{2cm}}

\hline
 \textbf{QC criteria} &  \textbf{QC criteria limit} &  \textbf{QC values} &  \textbf{QC status}\\
\hline   
1. Minimum Total Events per well 
&     $\ge$8000 total events        
&  `r format(min.events.per.well, scientific = FALSE)`  
&  `r min.events.per.well.status` \\


2. Assay Controls for Marker Assay 
\newline \hspace*{0.3cm}- Positive (Dx [10ng]) Marker Conc. 
\newline \hspace*{0.3cm}- NTC (H2O) Marker Conc. 
& \leavevmode\newline $\ge$10 copies/uL \newline $\le$0.1 copies/uL 
& \leavevmode\newline `r min.dx.marker.con.perwell.formatted` 
             \newline `r max.h2o.marker.con.perwell.formatted` 
& \leavevmode\newline `r min.dx.marker.con.perwell.status`    
             \newline `r max.h2o.marker.con.perwell.status` \\

3. Assay Controls for Albumin Assay 
\newline \hspace*{0.3cm}- Positive (MNC [5ng]) ALB Conc. 
\newline \hspace*{0.3cm}- NTC (H2O) ALB Conc.
& \leavevmode\newline $\ge$50 copies/uL \newline $\le$0.1 copies/uL 
& \leavevmode\newline `r min.mnc.alb.con.perwell.formatted` 
             \newline `r max.h2o.alb.con.perwell.formatted` 
& \leavevmode\newline `r min.mnc.alb.con.perwell.status`    
             \newline `r max.h2o.alb.con.perwell.status` \\

4. LPC concentration (copies/uL) & & `r sprintf("%#.2f", qc.concentration)` & \\

\hline

\textbf{Overall QC result} &     &  & `r final.qc.call` \\

\hline
\end{tabular}


# MRD calculation

```{r mrd_calculation, echo=FALSE, comment=""}
# cell contration at fullow up
fu.alb.info <- get_merged_info(c(params$fu.alb))
fu.alb.dilution.factor <- as.numeric(gsub("x$", "", strsplit(rownames(fu.alb.info), "_", fixed=TRUE)[[1]][4]))
fu.alb.concentration <- fu.alb.info[1, "concentration"] * fu.alb.dilution.factor
fu.cell.concentration <- fu.alb.info[1, "concentration"] * fu.alb.dilution.factor / 2
if (nrow(fu.alb.info) > 1) {
	print("warning, there are more than 1 albumin sample at follow up.")
}
# marker concentration at follow up
fu.marker.info <- get_merged_info(c(params$fu.marker))
# absolute mrd
absolute.mrd <- fu.marker.info[1, "concentration"] / fu.cell.concentration

writeLines("Values input:")
lineNum <- 0

if (is_manual_threshold){
        lineNum <- lineNum + 1
        writeLines(paste(lineNum, ". Using marker assay's intensities threshold ", manual_threshold, sep=""))
}

# diagnostic baseline
dx.baseline <- params$dx.baseline
if (is.na(dx.baseline)) {
	dx.alb.info <- get_merged_info(c(params$dx.alb))
	dx.cell.concentration <- dx.alb.info[1, "concentration"] / 2
	dx.marker.info <- get_merged_info(c(params$dx.marker))
	dx.baseline <- dx.marker.info[1, "concentration"] / dx.cell.concentration

	lineNum <- lineNum + 1
	writeLines(paste(lineNum, ". Using calculated diagnostic baseline.", sep=""))
} else {
	lineNum <- lineNum + 1
	writeLines(paste(lineNum, ". Using preset diagnostic baseline ", dx.baseline, sep=""))
}

lineNum <- lineNum + 1
writeLines(paste(lineNum, ". Using preset MNC Albumin concentration ", params$mnc.alb.concentration, sep=""))

lineNum <- lineNum + 1
writeLines(paste(lineNum, ". Using MNC dilution factor ", params$mnc.alb.dilutionX, sep=""))





# mrd in mnc
mnc.marker.info <- get_merged_info(c(params$mnc.marker))

if (mnc.marker.info[1, "positive"] == 0) {
	mnc.marker.info[1, "positive"] <- 1
	mnc.marker.info[1, "concentration"] <- -log(1-(mnc.marker.info[1, "positive"]/mnc.marker.info[1, "used.events"]))/0.00085
	lineNum <- lineNum + 1
	writeLines(paste(lineNum, ". No positive event of marker observed in MNC wells. 1 positive event assumed.", sep=""))
}

mnc.alb.info <- get_merged_info(c(params$mnc.alb))
mnc.alb.dilution.factor <- as.numeric(gsub("x$", "", strsplit(rownames(mnc.alb.info), "_", fixed=TRUE)[[1]][3]))

#background.level <- mnc.marker.info[1, "concentration"]*2/(mnc.alb.info[1, "concentration"] * mnc.alb.dilution.factor)
background.level <- mnc.marker.info[1, "concentration"]/(params$mnc.alb.concentration * params$mnc.alb.dilutionX / 2)

# total cell no.
total.cell.no <- ceiling(fu.cell.concentration * 20 * fu.marker.info[1, "replicates"])

# thresholds
ptv <- qbinom(0.9999, total.cell.no, background.level)/(20 * fu.marker.info[1, "replicates"])
ntv <- qbinom(0.95, total.cell.no, background.level)/(20 * fu.marker.info[1, "replicates"])


absolute.ptv <- ptv/fu.cell.concentration
absolute.ntv <- ntv/fu.cell.concentration





fu.marker.formatted <- sprintf("%#.2f", fu.marker.info[1, "concentration"])
fu.alb.concentration.formatted <- sprintf("%#.2f", fu.alb.concentration)
absolute.mrd.formatted <- formatC(absolute.mrd, format = "E", digits = 2)
absolute.ptv.formatted <- formatC(absolute.ptv, format = "E", digits = 2)
absolute.ntv.formatted <- formatC(absolute.ntv, format = "E", digits = 2)
background.level.formatted <- formatC(background.level, format = "E", digits = 2)

if (params$runType == "absolute") {


		
		
		if (absolute.mrd >= absolute.ptv) {
			mrd.call <- "POS"
		} else if (absolute.mrd < absolute.ntv) {
			mrd.call <- "NEG"
		} else {
			mrd.call <- "PNQ"
		}

		mrd.output.values <- c(fu.marker.formatted, 
					   		   fu.alb.concentration.formatted, 
					           absolute.mrd.formatted,  

					           absolute.ptv.formatted, 
					           absolute.ntv.formatted, 
			                   
					           background.level.formatted, 
					           total.cell.no,  
					           mrd.call)

		names(mrd.output.values) <- c("FU Marker (copies/uL)", 
		                              "FU ALB (copies/uL)", 
		                              "Absolute MRD", 

		                              "Absolute PTV",
		                              "Absolute NTV", 
			                          
		                              "MNC Baseline", 
		                              "Total Assessable GE Number", 
		                              "Call")


} 




mrd.output.values <- as.matrix(mrd.output.values)

colnames(mrd.output.values) <- "Values"

kable(mrd.output.values, caption = "Summary of MRD calculation.")



```

\vspace{2cm}
\begin{tabular}{|p{2cm}|p{4cm}|p{4cm}|p{4cm}|}
\hline
 & Name & Signature & Date\\
\hline
&&&\\
Run by&        `r params$run_by`     &  &    `r params$run_date`  \\
&&&\\
\hline
&&&\\
Reported by &  `r params$reported_by`   &  &   \\
&&&\\
\hline
&&&\\
Verified by &  `r params$verifier`  &  & \\
&&&\\
\hline
\end{tabular}



